\vspace{3cm} \Huge{Optimizing SasView models} \\ \Large{Enhancing performance by rewriting kernels from OpenCL/OpenMP to Futhark}

### 2017-11-07 report/draft.tex
qqq

\documentclass[11pt]{article}
\usepackage[a4paper, hmargin={2.8cm, 2.8cm}, vmargin={2.5cm, 2.5cm}]{geometry}
\usepackage{eso-pic} % \AddToShipoutPicture
\usepackage{graphicx} % \includegraphics
\usepackage{listings} % \includegraphics

@@

\section{Performance tests}

% description of built-in sasmodels comparison functionality
%% description of how comparison module has dictated the design of the futhark kernel (i.e. loading as much data as possible on init)

% display results

% discussion of futhark vs. python performance
%% where does the performance gains come from?

% discussion of futhark vs. opencl performance
%% where does the performance gains come from?
% show core_shell-whatever; and elaborate on loop flattening into maps
%%% MASSER AF REFERENCER TIL TROELS' PAPERS

% correctness

>>>>

\documentclass[11pt]{article}
\usepackage[a4paper, hmargin={2.8cm, 2.8cm}, vmargin={2.5cm, 2.5cm}]{geometry}
\usepackage{eso-pic} % \AddToShipoutPicture
\usepackage{graphicx} % \includegraphics
\usepackage{listings} % \includegraphics
\usepackage{pgfplots}

@@

\section{Method and performance results}
% description of built-in sasmodels comparison functionality
\sasmodels contains a comparison module called \texttt{compare.py}, which
lets the user run different calculations on the same dataset.
This module has been the basis of all my performance tests, and indeed also
the module that has let me design Futhark models in the first place.

\subsection{Development method}
The development of Futhark for \sasmodels has been largely test-driven.
For any given already existing model in the \sasmodels catalogue that I have
wished to implement in Futhark, I have started out by defining an equivalent
model written in Futhark, and then used the comparison module to compare the
 runs of the original model versus the Futhark implementation.

The comparison module displays the average execution time of each of the models,
which lets me track exactly how much faster (or slower) the Futhark model has
been running, compared to the original implementation.

Furthermore, and ultimately most important, the comparison module also
revealed the difference in results between the two models. 
As I have precisely zero knowledge on the field of Small Angle Scatterings, 
the only option for me regarding correctness of the models has been to take
the results of the already existing models as the "correct" results of the
computations. Consequently, any difference between my Futhark results and the
original results has been direct indicators of faulty implementations or bugs
somewhere along my work.
Therefore, I have been developing and debugging my solution by repeatedly
comparing it's results to that of the original implementation.

Conversely, I have chosen to accept test runs with exactly no difference
between the original and the Futhark implementation, as indicators of a
correct implementation. 
Test runs with thousands of data points defined as double-precision
with more than ten decimal points have still resulted in a total difference of
zero\footnote{see \ref{sec:correctness}} between the models.
This is what I base my assertion of a correctness on, as even the smallest 
mistakes along the calculations would surely have resulted in a very significant
accumulated error in the final results.

The test driven development has also lead to some of the biggest performance
optimizations in my design of kernelfut.py.
In example, the initial tests showed that the Futhark kernels were unreasonably
much slower than the already existing kernels. After some investigation I found
that this was because the original implementation buffered the input data in the
OpenCL context long time before the actual kernel was called in Python, whereas
the first versions of the Futhark kernel buffered the input data in the OpenCL
context as part of the \texttt{run\_kernel} function call.

\subsection{Futhark vs. Python performance}
I have tested Futhark vs. Python performance on two different models; 
the \texttt{line} model and the \texttt{broad\_peak} model.

\subsubsection{\texttt{line} model}
The \iq function for the line model is defined as $I(q) = A + B*q$, which is
equivalent to one of the most basic high school math functions, $f(x) = ax + b$.

\begin{figure}
    \begin{tikzpicture}
      \begin{axis}[
        mbarplot,
        title={1D performance},
        xlabel={number of datapoints},
        ylabel={Time in milliseconds},
        width=0.9\textwidth,
        height=3.5cm,
        symbolic x coords={100,500,2500,5000,10000},
        bar width=9pt,
        enlargelimits=0.15,
        ybar=10pt,% configures ‘bar shift’
        bar width=12pt,
        nodes near coords,
        legend style={legend pos=north west}
      ]
      \addplot plot coordinates {(100, 0.24 ) (500, 0.24 ) (2500, 0.26 ) (5000, 0.26 ) (10000, 0.28 )};
      \addplot plot coordinates {(100, 0.93 ) (500, 0.94 ) (2500, 0.97 ) (5000, 0.95 ) (10000, 0.97 )};
      \legend{Current, Futhark}
      \end{axis}
    \end{tikzpicture}
    \label{fig:line-graph}
    \caption{Comparison between Python and OpenCL performance for simple model}
\end{figure}

In figure \ref{fig:line-graph} we see that for one dimensional calculations,
both implementations keep a steady execution time. Futhark is consistently 
slower, but that can be explained by the overhead of using the Futhark kernels.

More interesting is it to look at the execution time of the 2D version of the 
experiment. For all \sasmodels 2D experiments, we are not just calculating \iq
for each q, but actually $\sum_{q \in Q} \sum_{q \in Q} Iq(q_x, q_y)$.
So if we have $n$ q, we need to perform $n^2$ calculations

n datapoints of q results in
having to calculate the scattering for q squared .
we have that we are actually 
¨

\subsubsection{Where does the performance gains come from?}

\subsection{Futhark vs. OpenCL performance}
\subsubsection{Where does the performance gains come from?}
% show core_shell-whatever; and elaborate on loop flattening into maps
%%% MASSER AF REFERENCER TIL TROELS' PAPERS

\subsection{Correctness of Futhark results}
